\appendix

\section{Appendix}

\subsection{Model from Robertson-Tessi and al. }
In this section, we will briefly describe the model from \cite*{Robertson-Tessi2015}.
The author describe an \acrfull{hca} model which is simulated on a grid with square lattice.
Each grid point can be empty or occupied by a cell.
Cell consumptions, energy production and acid secretion are computed by several functions and included into a PDE to determine substrates concentrations accross the microenvironment.

Cells will produce a target level of \acrshort{atp} using mostly aerobic pathways when oxygen is available. 
Oxygen consumption is described using a Michaelis-Menten function.
\begin{equation}
    \label{appendix:fo-robertson}
    f_O = - V_O { \frac{O}{ O + k_O} }
\end{equation}
with $V_O$ the maximum oxygen consumption rate, $O$ the extracellular oxygen concentration and $k_O$ the extracellular oxygen concentration at which oxygen consumption is half maximum. \\
Anaerobic pathways are an alternative way to produce \acrshort{atp} when cell requirement is not met using aerobic pathways.
Therefore, glucose consumption is increased to make up for the lack of \acrshort{atp}.
Glucose consumption is determined by a Michaelis-Menten equation with a modified term to reflect the increase of glucose consumption when oxygen is not available.
\begin{equation}
    \label{appendix:fg-robertson}
    f_G = - ( \frac{p_{G} A_0}{2} + \frac{27 f_O}{10} ) { \frac{G}{ G + k_G} }
\end{equation}
where $G$ is the extracellular glucose concentration, $k_G$ the half-maximum, $A_0$ the target level of \acrshort{atp} and $p_G$ is a coefficient to account for the increased glucose metabolism observed for many tumour cells.
For normal cells have $p_G = 1$, while for tumour cells $p_G>1$.
With higher values of $p_G$, cells consume more glucose to produce \acrshort{atp} than normal cells.
Effective \acrshort{atp} production is deduced from nutrient consumption.
\begin{equation}
    \label{appendix:fa-robertson}
    f_{A} = - ( 2 f_G + \frac{27 f_O}{5} ) \\
\end{equation}
The amount of glucose used through anaerobic pathways determine the quantity of protons secreted.
\begin{equation}
    \label{appendix:fh-robertson}
    f_{H^+} = k_H ( \frac{ 29 ( p_G V_O + f_O) }{5})
\end{equation}
where $k_H$ is a proton buffering coefficient.

Cell consumptions and productions are used to dermine the concentration of a molecule accross the microenvironment.
The concentration $C$ of a molecule $x$ is described by the equation :
\begin{equation}
    \label{appendix:pde-robertson}
    \frac{\partial C}{\partial t} = D \nabla^2 C + f(C,p) 
\end{equation}
where $D$ is a diffusion coefficient, $f$ is a function accounting for the impact of cell with parameters $p$ on the local substrate concentration $C$.
Diffusing substrates are oxygen ($O$), glucose ($G$) and protons ($H$).
Functions $f_O$, $f_G$ and $f_H$ are used for the $f$ term in equation \ref*{appendix:pde-robertson}, $f_A$ is used to drive the cells behaviour in the cellular automaton.